The Revisited Hidden Weight Bit Function

### 2025-08-29 paper.tex
Minor
Minors

\newcommand{\AI}{\mathsf{AI}}
\newcommand{\Bfn}[1]{\mathcal{B}_{#1}}
\newcommand{\BN}{\mathcal{B}_n}
\newcommand{\C}{\mathbb{C}}
\newcommand{\dd}{\,\mathrm d}
\newcommand{\degg}{\mathsf{deg}}
\newcommand{\dis}{\mathsf{c_1}}
\newcommand{\Dkna}[3]{\mathsf{D}_{#2,#1}(a)}
\newcommand{\Dtkna}[4]{\mathsf{D}_{#1,#2,#3}(#4)}
\newcommand{\ee}{\mathrm e\mkern1mu}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\Ekn}[2]{\mathsf{E}_{#1,#2}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\FAI}{\mathsf{FAI}}
\newcommand{\hd}{\mathsf{d_H}}
\newcommand{\hwbf}{\textsf{HWBF}}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\ii}{\mathrm i\mkern1mu}
\newcommand{\kraw}[3]{\mathsf{K}_{#1}(#2,#3)}
\newcommand{\myurl}[1]{\footnote{\texttt{\url{#1}}}}
\newcommand{\NL}{\mathsf{NL}}
\newcommand{\phikn}[2]{\varphi_{#1,#2}}
\newcommand{\w}{\mathsf{w_H}}
\newcommand{\WD}[2]{\mathcal{WD}^{#1}_{#2}}
\newcommand{\wt}[1]{\mathcal W_{#1}} 
\newcommand{\wtk}[2]{\mathcal{W}_{#1,#2}}
\newcommand{\Z}{\mathbb{Z}}

@@

\author{%
    Pierrick Méaux\inst{1}\orcidID{0000-0001-5733-4341} \and
    Tim Seuré\inst{1}\orcidID{0000-0003-4200-7553} \and %Corresponding author
    Deng Tang\inst{2}\orcidID{0000-0002-8373-9200}
}
\authorrunning{P. Méaux et al.}

@@

    	\keywords{Boolean functions, HWBF, Nonlinearity.}
    \end{abstract}

@@

    These various results on the algebraic properties of the revisited \hwbf{} suggest that it possesses strong resistance against standard attacks. Furthermore, its high nonlinearity makes it a well-suited candidate for use as a filter function in contexts such as filtered \textsf{LFSR}s or in homomorphically-friendly schemes like \textsf{FLIP} and \textsf{FiLIP}. For example, the best current \textsf{FiLIP} filters in \textsf{HHE}~\cite{eprint:AGHM24,CCS:CDPP22,CIC:MeaParPei24} are \textsf{XOR}-threshold functions—the sum of a $k$-variable linear function and an $m$-variable threshold function (generalizing majority).  
    By Proposition $7$ in~\cite{IEEE:CarMea21}, their nonlinearity is at most:
    \[
    2^{k+m-1} - 2^k\cdot M, \quad \text{where } M = \begin{cases}
    \binom{m-1}{\frac{m-1}{2}} & \text{if } m \text{ is odd}, \\
    \frac{1}{2} \binom{m}{\frac{m}{2}} & \text{if } m \text{ is even}.
    \end{cases}
    \]
    This is similar to the majority function, whereas the revisited \hwbf{} achieves higher nonlinearity for the same size, offering better resistance to correlation-like attacks with fewer variables.
    \begin{table}
    	\scriptsize
    	\centering
        \caption{Lowest possible values of $(d, e)$ for different values of $n$.}
    	\label{table:FAI}
    	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
            \hline
            $n$&$4$&$6$&$8$&$10$&$12$&$14$&$16$&$18$\\
            \hline
            $(d,e)$&$(1,2)$&$(1,3)$&$(1,5)$&$(1,7)$&$(1,9)$&$(1,11)$&$(1,13)$&$(1,15)$\\
            &&&$(2,4)$&$(2,6)$&$(2,8)$&$(2,10)$&$(2,12)$&$(2,14)$\\
            &&&&$(3,5)$&$(3,6)$&$(3,9)$&$(3,11)$&$(3,13)$\\
            &&&&&&&$(4,9)$&$(4,11)$\\
            &&&&&&&&$(5,10)$\\
            \hline
    	\end{tabular}
    \end{table}

@@

    Our approach utilizes generating functions and Cauchy’s estimate to establish lower bounds on the nonlinearity of weightwise quadratic functions. This naturally raises the question: can the employed techniques, particularly the use of Cauchy’s estimate, be extended to other families of Boolean functions? Specifically, this approach appears feasible for other weightwise-degree-$d$ functions $f$ where the $f_k$ are direct sums. Investigating this could potentially lead to the discovery of new functions with comparable computational costs and even better cryptographic properties.

@@

        Pierrick Méaux was funded by the European Research Council (\textsf{ERC}) under the Advanced Grant program (reference number: 787390). Tim Seuré acknowledges the support of the Luxembourgish “Fonds National de la Recherche” (\textsf{FNR}) through an Individual Grant (reference number: 17936291). The work of Deng Tang was supported in part by the National Natural Science Foundation of China (\textsf{NSFC}, reference number: 62272303). We would also like to thank Claude Carlet for discussing the function introduced in this work with Deng Tang some years ago.
    \end{credits}

>>>>

\newcommand{\AI}{\mathsf{AI}}
\newcommand{\Bfn}[1]{\mathcal{B}_{#1}}
\newcommand{\BN}{\mathcal{B}_n}
\newcommand{\C}{\mathbb{C}}
\newcommand{\dd}{\,\mathrm d}
\newcommand{\degg}{\mathsf{deg}}
\newcommand{\dis}{\mathsf{c_1}}
\newcommand{\Dkna}[3]{\mathsf{D}_{#2,#1}(a)}
\newcommand{\Dtkna}[4]{\mathsf{D}_{#1,#2,#3}(#4)}
\newcommand{\ee}{\mathrm e\mkern1mu}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\Ekn}[2]{\mathsf{E}_{#1,#2}}
\newcommand{\etal}{\textsl{et al.}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\FAI}{\mathsf{FAI}}
\newcommand{\hd}{\mathsf{d_H}}
\newcommand{\hwbf}{\textsf{HWBF}}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\ii}{\mathrm i\mkern1mu}
\newcommand{\kraw}[3]{\mathsf{K}_{#1}(#2,#3)}
\newcommand{\myurl}[1]{\footnote{\texttt{\url{#1}}}}
\newcommand{\NL}{\mathsf{NL}}
\newcommand{\phikn}[2]{\varphi_{#1,#2}}
\newcommand{\w}{\mathsf{w_H}}
\newcommand{\WD}[2]{\mathcal{WD}^{#1}_{#2}}
\newcommand{\wt}[1]{\mathcal W_{#1}} 
\newcommand{\wtk}[2]{\mathcal{W}_{#1,#2}}
\newcommand{\Z}{\mathbb{Z}}

@@

\author{%
    Pierrick Méaux\inst{1}\orcidID{0000-0001-5733-4341} \and
    Tim Seuré\inst{1}\orcidID{0000-0003-4200-7553} \and %Corresponding author
    Deng Tang\inst{2}\orcidID{0000-0002-8373-9200}
}
\authorrunning{P. Méaux \etal}

@@

    	\keywords{Boolean functions \and HWBF \and Nonlinearity.}
    \end{abstract}

@@

    These various results on the algebraic properties of the revisited \hwbf{} suggest that it possesses strong resistance against standard attacks. Furthermore, its high nonlinearity makes it a well-suited candidate for use as a filter function in contexts such as filtered \textsf{LFSR}s or in homomorphically-friendly schemes like \textsf{FLIP} and \textsf{FiLIP}. For example, the best current \textsf{FiLIP} filters in \textsf{HHE}~\cite{eprint:AGHM24,CCS:CDPP22,CIC:MeaParPei24} are \textsf{XOR}-threshold functions—the sum of a $k$-variable linear function and an $m$-variable threshold function (generalizing majority). By Proposition $7$ in~\cite{IEEE:CarMea21}, their nonlinearity is at most $2^{k+m-1}-2^k\cdot M$, where $M=\binom{m-1}{\frac{m-1}{2}}$ if $M$ is odd, and $M=(1/2)\cdot\binom{m}{\frac{m}{2}}$ if $M$ is even. This is similar to the majority function, whereas the revisited \hwbf{} achieves higher nonlinearity for the same size, offering better resistance to correlation-like attacks with fewer variables.
    \begin{table}
    	\scriptsize
    	\centering
        \caption{Lowest possible values of $(d, e)$ for different values of $n$.}
    	\label{table:FAI}
    	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
            \hline
            $n$&$4$&$6$&$8$&$10$&$12$&$14$&$16$&$18$\\
            \hline
            $(d,e)$&$(1,2)$&$(1,3)$&$(1,5)$&$(1,7)$&$(1,9)$&$(1,11)$&$(1,13)$&$(1,15)$\\
            &&&$(2,4)$&$(2,6)$&$(2,8)$&$(2,10)$&$(2,12)$&$(2,14)$\\
            &&&&$(3,5)$&$(3,6)$&$(3,9)$&$(3,11)$&$(3,13)$\\
            &&&&&&&$(4,9)$&$(4,11)$\\
            &&&&&&&&$(5,10)$\\
            \hline
    	\end{tabular}
    \end{table}

@@

    Our approach utilizes generating functions and Cauchy's estimate to establish lower bounds on the nonlinearity of weightwise quadratic functions. This naturally raises the question: can the employed techniques, particularly the use of Cauchy's estimate, be extended to other families of Boolean functions? Specifically, this approach appears feasible for other weightwise-degree-$d$ functions $f$ where the $f_k$ are direct sums. Investigating this could potentially lead to the discovery of new functions with comparable computational costs and even better cryptographic properties.

@@

        Pierrick Méaux was funded by the European Research Council (\textsf{ERC}) under the Advanced Grant program (reference number: 787390). Tim Seuré acknowledges the support of the Luxembourgish “Fonds National de la Recherche” (\textsf{FNR}) through an Individual Grant (reference number: 17936291). The work of Deng Tang was supported in part by the National Natural Science Foundation of China (\textsf{NSFC}, reference number: 62272303). We would also like to thank Claude Carlet for discussing the function introduced in this work with Deng Tang some years ago.

        \subsubsection{\discintname}
        The authors have no competing interests to declare
        that are relevant to the content of this article.
    \end{credits}
